Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-10-23 other/confnoncoherent/paper.tex
minor

\begin{proposition}\label{prop:contgg2}
Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.  Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let,
\[
G(x) =  \expect R \cos\sfracpart{x + \Phi}.
\]
Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
\end{proposition}

>>>>

\begin{proposition}\label{prop:contgg2}
Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.  Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let,
\[
G(x) =  \expect R \cos\sfracpart{x + \Phi}.
\]
Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
\end{proposition}

### 2012-10-23 paper.tex
minor

\begin{lemma}\label{lem:expRvlamphi}
$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
\end{lemma}
\begin{IEEEproof}
In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.  An analogous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.  Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, 
\begin{align}
W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi, \label{eq:Winfsum}
\end{align}
where $g$ is defined in the statement of Theorem~\ref{thm:normality}
Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
\begin{align*}
\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
\end{align*}
where, for $k = 1, \dots, M-1$,
\begin{equation}\label{eq:ak}
a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
\end{equation}
\[
b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
\]
\begin{equation}\label{eq:ck}
c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
\end{equation}
Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}, it follows that $c_k(\lambda)$ converges to 
\[
s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
\]
as $\lambda$ goes to zero from above.  We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.  The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.

@@

\begin{lemma}\label{lem:convdistGLdash}
The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
\end{lemma}
\begin{IEEEproof}
Observe that $\sqrt{L} R_L(0) = C_L + D_L$ where
\[
C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i), \qquad D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
\]
From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
\[
p A_1 = p \expect R_1^2 \sin^2(\Phi_1).
\]
Similarly, the distribution of $D_L$ converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
\[
d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
\]
The lemma holds because $C_L$ and $D_L$ are independent. 
\end{IEEEproof}

@@

\begin{lemma}\label{lem:XL} 
The distribution of 
\[
X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
\] 
converges to the normal with zero mean and covariance 
\[
pB_1 + d B_2
\] 
as $L \rightarrow\infty$.
\end{lemma}
\begin{IEEEproof}
Observe that $X_L = C_L + D_L$ where
\[
C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
\]
\[
D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ),
\]
where $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.  From the standard central limit theorem the distribution of $C_L$ converges to the normal with zero mean and variance
\[
p B_1 = p \expect R_1^2 \cos^2(\Phi_1) - p h_1^2(0) =  p \expect R_1^2 \cos^2(\Phi_1) - p 
\]
since $h_1(0) = 1$.  Similarly the distribution of $D_L$ converges to the normal with zero mean and variance
\[
d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0).
\]
The lemma follows since $C_L$ and $D_L$ are independent.
\end{IEEEproof}

@@

>>>>

\begin{lemma}\label{lem:expRvlamphi}
$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
\end{lemma}
\begin{IEEEproof}
In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.  An analogous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.  Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, 
\begin{align}
W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi, \label{eq:Winfsum}
\end{align}
where $g$ is defined in the statement of Theorem~\ref{thm:normality}.
Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
\begin{align*}
\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
\end{align*}
where, for $k = 1, \dots, M-1$,
\begin{equation}\label{eq:ak}
a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
\end{equation}
\[
b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
\]
\begin{equation}\label{eq:ck}
c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
\end{equation}
Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}, it follows that $c_k(\lambda)$ converges to 
\[
s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
\]
as $\lambda$ goes to zero from above.  We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.  The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.

@@

\begin{lemma}\label{lem:convdistGLdash}
The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ as $L\rightarrow\infty$.
\end{lemma}
\begin{IEEEproof}
Observe that $\sqrt{L} R_L(0) = C_L + D_L$ where
\[
C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i), \qquad D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
\]
From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
\[
p A_1 = p \expect R_1^2 \sin^2(\Phi_1).
\]
Similarly, the distribution of $D_L$ converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
\[
d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
\]
The lemma holds because $C_L$ and $D_L$ are independent. 
\end{IEEEproof}

@@

\begin{lemma}\label{lem:XL} 
The distribution of 
\[
X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
\] 
converges, as $L \rightarrow\infty$,  to the normal with zero mean and covariance $pB_1 + d B_2.$
\end{lemma}
\begin{IEEEproof}
Observe that $X_L = C_L + D_L$ where
\[
C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
\]
\[
D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ),
\]
where $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.  From the standard central limit theorem the distribution of $C_L$ converges to the normal with zero mean and variance
\[
p B_1 = p \expect R_1^2 \cos^2(\Phi_1) - p h_1^2(0) =  p \expect R_1^2 \cos^2(\Phi_1) - p 
\]
since $h_1(0) = 1$.  Similarly the distribution of $D_L$ converges to the normal with zero mean and variance
\[
d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0).
\]
The lemma follows since $C_L$ and $D_L$ are independent.
\end{IEEEproof}

@@

\begin{IEEEproof}
We first prove the result for $h_1$.  Write
\begin{align*}
h_1(x) &= \int_{-\pi}^{\pi} \int_{0}^\infty r \cos(\phi + x) f(r,\phi) dr d\phi \\
&= \int_{-\pi}^{\pi}  \cos(\phi + x) g(\phi) d\phi,
\end{align*}
where $f(r,\phi)$ is the joint pdf of $R$ and $\Phi$, and 
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
\]
Since the pdf of $X$ is circularly symmetric, $f(r,\phi)$ and $g(\phi)$ are even in $\phi$ on $[-\pi,\pi]$.

Write $\cos(\phi + x)$ as a sum of even and odd parts,
\[
\cos(\phi + x) = \cos(\phi)\cos(x) - \sin(\phi)\sin(x).
\]
Now
\[
h_1(x) = \cos(x) \int_{-\pi}^{\pi} \cos(\phi) g(\phi) d\phi
\]
since $\int_{-\pi}^{\pi} \sin(\phi) g(\phi) d\phi = 0$ because $\sin(\phi)g(\phi)$ is odd.  Now 
\[
K = \int_{-\pi}^{0} \cos(\phi) g(\phi) d\phi > 0
\] 
because $g(\phi) > 0$ is increasing on $[-\pi,0]$ and $\cos(\phi)$ is odd and increasing on $[-\pi,0]$.  Thus $h_1(x) = 2K \cos(x)$ is uniquely maximised at zero because $\cos(x)$ is.
\end{IEEEproof}

\begin{lemma}
Under the assumptions in Proposition~\ref{prop:contgg2} let $f_Z$ be the pdf of $\abs{W}$.  Then $z^{-1} f_Z(z)$ is differentiable and nonincreasing in $z > 0$ and the function $g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr$ is symmetric and unimodal with mode at $\phi = 0$.
\end{lemma}
We have
\[
f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{\sqrt{x^2 + y^2}}.
\]
Put $z = \sqrt{x^2 + y^2}$ and, since $X$ is circularly symmetric,
\[
f_{X,Y}(x,y) = f_{X,Y}(z,0) = \frac{f_Z(z)}{z}.
\]
Since $f_{X,Y}(z,0)$ is differentiable and nonincreasing in $z = \sqrt{x^2 + y^2}$ (by assumption in the proposition statement) we have that $z^{-1} f_Z(z)$ is differentiable and nonincreasing in $z > 0$.  So the derivative $\frac{d}{dz} \big( z^{-1} f_Z(z) \big) \geq 0$ for all $z > 0$.   Using~\eqref{eq:pdfRPhi},
\[
g(\phi) = \int_{0}^\infty r^2 \frac{f_Z(z)}{2\pi z } dr
\]
where $z = \sqrt{r^2 - 2r\cos\phi + 1}$.  Differentiating with respect to $\phi$,
\[
g^\prime(\phi) = \sin(\phi) \int_{0}^\infty \frac{r^3}{\pi} \frac{d}{dz}\left( \frac{f_Z(z)}{z} \right) dr
\]
and since the term inside the integral is always positive,
